Technical Report

### 2018-02-23 fully-abstract-comp/technical-report.tex
.
Fix three technical issues in the world satisfaction that I just discussed with Lau.
Add a missing projection.
Added the rest of the memory satisfaction defintion.
removed the other value relation and xjmp. Made the format A4 as no definitions are too long for this.
updated the world sections. Hopefully, everything necessary for the new worlds is there (except a solution for the recursive domain equation).

\documentclass[a3paper]{article}
\usepackage[svgnames]{xcolor}
\usepackage[a3paper,margin=0.7in]{geometry}
\usepackage[utf8]{inputenc}

@@

% \dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
% \lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
% \lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \Caps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    \scbnf \in &\SealableCaps&\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    \cbnf \in&\Caps& \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}{\sourcecolor{} \; \mid \retptrd(\sigma,(\basebnf,\aendbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\ 
    &\rbnf & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus \{\halted\}
  \end{array}
\]
%\lau{06-09-2017: Should seals also have the option to be linear?}
%\lau{13-09-2017: For now I have added linearity to the seals.}
%\lau{15-12-2017: Removed linearity from seal - I could not come up with a single good reason to have it as we have them here.}
\dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
\lau{28-09-2017: Why would you like this? If we want to add this, we will have to be careful, because an old code-return-pointer may fit perfectly well with our new data-return-pointer, so in this case it should not fail. We added the seal to the return pointers, so we had some way of figuring out whether two return pointers would actually fit together.}
\dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
\lau{28-09-2017: At the moment, the only part of the stack that is stored in the stack frame is the private part. Specifically it is the part from the base to the current address of the stack pointer \emph{in $r_\stk$}. This means that if someone saves a stack pointer, so it is not included in the call, then they can still use it because that part of the stack was not encapsulated in a stack frame. Makes sense?}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\rbnf$ defines the finite set $\RegName$. 
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

@@

\noindent\textbf{xjmp}\\
\lau{22-09-2017: Can we allow the callee to return any stack pointer? At the moment, it is not possible to create an empty stack pointer which may be a saving grace. If we at some point change this, so you can split capabilities into empty ones, then we need to consider whether it is safe for callees to return empty stacks (I think it is okay as long as we always have some stack usage).}
\lau{28-09-2017: For now, the jump to a return pointer does the stack end point check. We could possibly move this to the proposed semantic condition as well.}
\begin{align*}
  \sem{\txjmp{r_1}{r_2}}(\Phi) = &
                                   \begin{cases}
                                     \arraycolsep=0pt
                                     \begin{array}[t]{rl}
                                       \Phi & \updReg{r_1}{w_1}\\
                                            & \updReg{r_2}{w_2}\\
                                            & \updReg{\pcreg}{c_\code}\\
                                            & \updReg{\rdata}{c_\data}
                                     \end{array} &
                                     \begin{array}[t]{l}
                                       \Phi(r_1) = \sealed{\sigma_1,c_\code} \tand \\
                                       \Phi(r_2) = \sealed{\sigma_2,c_\data} \tand \\
                                       \sigma_1 = \sigma_2 \tand \\
                                       \nonExec{c_\data} \tand \\
                                       w_1 = \linCons{\Phi(r_1)} \tand \\
                                       w_2 = \linCons{\Phi(r_2)}
                                     \end{array}
                                     \\&\\
                                     \sourcecolor\left.
                                       \arraycolsep=0pt
                                       \begin{array}[t]{rl}
                                         \Phi' & \updReg{r_2}{0} \\
                                               & \updReg{\pcreg}{\opc}\\
                                               & \updReg{\rdata}{0}\\
                                               & \updReg{\rstk}{c_\stk}\\
                                               & \updReg{\rtmp{1}}{0}\\
                                               & \updReg{\rtmp{2}}{0}
                                       \end{array}\right.&
                                     \sourcecolor\left.
                                       \begin{array}[t]{l}
                                         \Phi(r_1) = \retptrc(\sigma_1,\_) \tand \\
                                         \Phi(r_2) = \retptrd(\sigma_2,(\aaddr_\stk,\eaddr_{\stk,\priv})) \tand \\
                                         \sigma_1 = \sigma_2 \tand\\
                                         \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
                                         \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
                                         \stkf = (\opc,\ms_{\stk,\priv}) \tand \\
                                         \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
                                         \eaddr_{\stk} + 1 = \aaddr_\stk \tand\\
                                         c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},\aaddr_\stk} \tand\\
                                         \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
                                       \end{array}\right.
                                     \\
                                     \failed & \totherwise
                                   \end{cases}
\end{align*}

@@

                                                        &\updReg{\rretc}{\retptrc(\sigma_?,(\perm,\lin,\baddr,\eaddr,\aaddr+\calllen))}\\

@@

\section{Logical Relation}
In the following definitions, \src{blue} is used to indicate values related to the source machine. This is unlike previous definitions, where \src{blue} was used to indicate source language specific parts of definitions.
\subsection{Worlds}
\[
  \World = \Worldh \times \Worlds \times \Worldfs
\]
with $\Worlds$, $\Worldh$, and $\Region$ defined as follows
\[
  \Worldh = \RegionName \fun (\Regions + \Regionh)
\]
and
\[
  \Worlds = \RegionName \fun (\Regions \times \SealableCaps)
\]
and
\[
  \Worldfs = \RegionName \fun \Regions
\]
where $\RegionName = \nats$.
\[
  \Regionh = 
    \{\pure \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \URel{\MemSeg^2})
\]
and
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \monnefun \URel{\MemSeg^2}) \uplus \\
    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \monnefun \URel{\MemSeg^2})\uplus \\ 
    \{\revoked\}
  \end{array} \right.
\]
where $\spatial$ and $\spatialo$ are regions governing segments of memory addressed by linear capabilities. $\spatialo$ signifies that this region is addressable. $\spatial$ signifies that the region is not owned and can thus not be addressed. At the same time it signifies that if something else addresses it, it is a $\linear$ capability. Finally, $\pure$ signifies that the region is only addressed by non-linear capabilities. Notice that no region allows for both linear and non-linear capabilities to address it.
\lau{26-20-2017: We may also want to build the infrastructure for the stack more explicitly into the worlds (but maybe we can handle it in the memory satisfaction).}
  \[
   \Rels = \{(\phi_\pub, \phi) \in \powerset{\State^2}\times \powerset{\State^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
  \]

Disjoint union of worlds. Joins together two alike worlds with strictly disjoint ownership over $\spatialo$-regions.
\[
  W_1 \oplus W_2 = W
  \begin{array}[t]{l}
    \text{if $\dom(W) = \dom(W_1) = \dom(W_2) \tand$} \\
    \forall r \in \dom(W) \ldotp W(r) = W_1(r) \oplus W_2(r)
  \end{array}
\]
$\oplus$ on regions is defined as follows
\begin{align*}
  (\pure,\var{sm}) \oplus (\pure,\var{sm}) =  & \; (\pure,\var{sm}) \\
  (\spatial,\var{sm}) \oplus (\spatial,\var{sm}) =  & \; (\spatial,\var{sm}) \\
  (\spatialo,\var{sm}) \oplus (\spatial,\var{sm}) = & \; (\spatial,\var{sm}) \oplus (\spatialo,\var{sm})\\
                                           =  & \; (\spatialo,\var{sm})
\end{align*}
and for all other cases $\oplus$ is undefined. Specifically, $\oplus$ is not defined when both sides are a $\spatialo$-region. It is further not defined if the two sides do not agree on region type or state machine. Here $\var{sm}$ is the state machine of the region.

\subsection{Future world}
\subsubsection{\color{DarkGreen} first thoughts on change to future world relation}
{\color{DarkGreen}
The public part of the future world relation stays the same:
\begin{mathpar}
  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \pubft W(r) }
            { W' \future W }
\end{mathpar}
The private future world relation becomes:
\begin{mathpar}
  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \privft W(r) \\  \exists r \in \dom(W') \setminus \dom(m(W))\ldotp W'(r) \text{ is non-empty and stack region}}
            { W' \future W }
\end{mathpar}
where we take "is non-empty and stack region" to mean the following:

@@

\begin{definition}
  We say a region $(k,s,\phi_\pub,\phi,H)$ is a stack region on a machine with starting stack domain $[\stkb,\eaddr]$ iff
\[
  \begin{array}{l}
    \forall n, \hat{W},s'\ldotp \forall \npair{\stpair{\ms}{\ms}} \in H \; s' \; \hat{W} \ldotp \dom(\src{\ms_S}) \subseteq [\stkb,\eaddr]  \wedge \dom(\ms_T) \subseteq [\stkb,\eaddr]
  \end{array}
\]
\end{definition}
}

For both $\privft$ and $\pubft$ we have
\begin{mathpar}
  \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \future W(r) }
            { W' \future W }
\end{mathpar}
%
The private future regions allow transitions in the private part of the transition system. They also allow $\spatial$ regions to become $\revoked$:
\begin{mathpar}
  \inferrule{ }
  { \revoked \privft (\spatial,\_,\_,\_,\_)}
\and
  \inferrule{ }
            { \revoked \privft \revoked }
\and
  \inferrule{ (s,s') \in \phi }
            { (v,s',\phi_\pub,\phi,H) \privft (v,s,\phi_\pub,\phi,H)  }
\end{mathpar}
%
Public future regions are allowed to transition according to the public part of their transition system.
\begin{mathpar}
    \inferrule{ }
            { \revoked \pubft \revoked }
\and
  \inferrule{ (s,s') \in \phi_\pub }
            { (v,s',\phi_\pub,\phi,H) \pubft (v,s,\phi_\pub,\phi,H)  }
\end{mathpar}

\subsection{Memory satisfaction}
{\color{DarkGreen}
Memory satisfaction for new worlds:
\[
  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
      \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \wedge\\
      \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}\ldotp \\
      \quad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \wedge \\
      \quad \memSatStack{\stk,\ms_\var{T,stack}}{W} \wedge \\
      \quad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W} \wedge \\
      \quad \memSatHeap{\ms_S,\ms_\var{T,heap}}{W}
    \end{array}
  \right.
\]

\[
  \memSatStack{\stk,\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      W_\var{stack} = \pi_2(W) \wedge \\
      \stk = (\opc_0,\ms_0), \dots (\opc_m,\ms_m) \wedge \\
      \exists R_\ms : \dom(W_\var{stack}) \fun \MemSeg \times \SealableCaps \times \MemSeg \ldotp \\
      \quad \ms_T = \biguplus_{r \in \dom(W_\var{stack})} \pi_3(R_\ms(r)) \wedge \\
      \quad \\
      \quad \exists R_W : \dom(W_\var{stack}) \fun World \ldotp \\
      \qquad \forall r \in \dom(W_\var{stack}), n' < n \ldotp \\
      \qquad \quad \npair[n']{(\pi_1(R_\ms(r)),\pi_3(R_\ms(R))} \in W_\var{stack}(r).H \; W_\var{stack}(r).s \; \xi^{-1}(R_W(r)) \\
      \qquad \quad \pi_2(R_\ms(r)) = W_\var{stack}(r).\opc

    \end{array}
  \right.
\]

\[
  \memSatFStack{ms_\stk,\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      \dots
    \end{array}
  \right.
\]

\[
  \memSatHeap{ms_\stk,\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      \dots
    \end{array}
  \right.
\]

@@

\[
  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      \exists R_\ms : \dom(W) \fun \MemSeg^2 \ldotp \\
      \quad \ms_T = \biguplus_{r \in \dom(W)} \pi_2(R_\ms(r)) \wedge \\
      \quad \stk = (\_,\ms_{\stk,0}) :: \dots :: (\_,\ms_{\stk,m}) \wedge \\
      \quad \ms_S \uplus \ms_\stk \uplus \ms_{\stk,0} \uplus \dots \uplus \ms_{\stk,m} = \biguplus_{r \in \dom(W)} \pi_1(R_\ms(r)) \wedge \\
      \quad \exists R_w : \dom(W) \fun \World \ldotp\\
      \qquad W = \bigoplus_{r \in \dom(W)} R_w(r) \wedge \\
      \qquad \forall r \in \dom(W) , n' < n \ldotp \\
      \quad \qquad \npair[n']{R_\ms(r)} \in W(r).H \; (W(r).s) \; (\xi^{-1} (R_w (r))) 
    \end{array}
  \right.
\]
In the definition of memory satisfaction, $R_\ms$ creates a partitioning of the two related memories. As the source language is a special interpretation of the target language, part of the target memory needs to relate to the stack in the source language. We also need to split up the world which is done by $R_W$.We split up the world in order to make sure that the linearity constraints imposed by the world are satisfied.
\lau{30-10-2017: It is not clear to me whether this definition should also talk about $\opc$.}
\dominique{4-12-2017: Regarding the question whether $\opc$ should be mentioned in the memory satisfaction relation.
  The relevant intuitive aspect of the proof here, I think, is that on a return, the source semantics restores opc from the copy kept in the stack, while the target semantics restores it from the copy in the code part of the return pointer.
  We will need to prove (somewhere) that these two correspond, which means we will need to track of the invariant that guarantees this somewhere (perhaps here?).
  I suspect the relevant invariants are (1) that any executable pointer signed with $\sealAss{a}$ must point to $a + \calllen$ and (2) that anything related to a $\rretc$ must be executable and sealed.
  Perhaps this invariant is easiest to track in (1) the general requirement on the $H_\sigma$ and (2) in the value relation.
  This suggests nothing should be required about opc here?
  To be confirmed in the proof of the fundamental theorem.
}

>>>>

\documentclass[a4paper]{article}
\usepackage[svgnames]{xcolor}
\usepackage[a4paper,margin=0.7in]{geometry}
\usepackage[utf8]{inputenc}

@@

% \dominique{7-9-2017: Should we keep local/global and linear/non-linear?  I was thinking just the latter would suffice...}
% \lau{07-09-2017: The local/global mechanism still has some value as it now can be used for the purpose of arguments that cannot be stored by the receiver (if you are willing to ``pay'' the clearing cost). I do, however, not think that it is strictly necessary here to make things work.}
% \lau{12-09-2017: I will remove the local capabilities today for the reasons we have discussed (they add to the complexity and they seem to be of little to no use without a clear memory instruction).}
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \Caps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    \scbnf \in &\SealableCaps&\defbnf & ((\permbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\sealbasebnf,\sealendbnf,\sigma}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    \cbnf \in&\Caps& \defbnf &  \scbnf \mid \sealed{\sigma,\scbnf}{\sourcecolor{} \; \mid \retptrd(\sigma,(\basebnf,\aendbnf)) \mid \retptrc(\sigma,(\permbnf,\linbnf,\basebnf,\aendbnf,\addrbnf))}\\ 
    &\rbnf & \defbnf &\pcreg \mid \rretd \mid \rretc \mid \rstk \mid \rdata \mid \rtmp{1} \mid \rtmp{2} \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor \sstkframe \in} & {\sourcecolor \StkFrame} & {\sourcecolor \defeq} & {\sourcecolor \Caps \times \MemSeg}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus \{\halted\}
  \end{array}
\]
%\lau{06-09-2017: Should seals also have the option to be linear?}
%\lau{13-09-2017: For now I have added linearity to the seals.}
%\lau{15-12-2017: Removed linearity from seal - I could not come up with a single good reason to have it as we have them here.}
\dominique{27-9-2017: I think we should have versions of $\retptrd$ and $\retptrc$ for non-top stack frames and define that invoking them with the current stack pointer will fail.}
\lau{28-09-2017: Why would you like this? If we want to add this, we will have to be careful, because an old code-return-pointer may fit perfectly well with our new data-return-pointer, so in this case it should not fail. We added the seal to the return pointers, so we had some way of figuring out whether two return pointers would actually fit together.}
\dominique{27-9-2017: Likewise, I think we need to have a source-language representation for a pointer into one of the non-top stack frames, as we may need to be able to back-translate adversary code that reads/writes in those.}
\lau{28-09-2017: At the moment, the only part of the stack that is stored in the stack frame is the private part. Specifically it is the part from the base to the current address of the stack pointer \emph{in $r_\stk$}. This means that if someone saves a stack pointer, so it is not included in the call, then they can still use it because that part of the stack was not encapsulated in a stack frame. Makes sense?}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\rbnf$ defines the finite set $\RegName$. 
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

We now define

@@

\noindent\textbf{xjmp}\\
\lau{22-09-2017: Can we allow the callee to return any stack pointer? At the moment, it is not possible to create an empty stack pointer which may be a saving grace. If we at some point change this, so you can split capabilities into empty ones, then we need to consider whether it is safe for callees to return empty stacks (I think it is okay as long as we always have some stack usage).}
\lau{28-09-2017: For now, the jump to a return pointer does the stack end point check. We could possibly move this to the proposed semantic condition as well.}
% \begin{align*}
%   \sem{\txjmp{r_1}{r_2}}(\Phi) = &
%                                    \begin{cases}
%                                      \arraycolsep=0pt
%                                      \begin{array}[t]{rl}
%                                        \Phi & \updReg{r_1}{w_1}\\
%                                             & \updReg{r_2}{w_2}\\
%                                             & \updReg{\pcreg}{c_\code}\\
%                                             & \updReg{\rdata}{c_\data}
%                                      \end{array} &
%                                      \begin{array}[t]{l}
%                                        \Phi(r_1) = \sealed{\sigma_1,c_\code} \tand \\
%                                        \Phi(r_2) = \sealed{\sigma_2,c_\data} \tand \\
%                                        \sigma_1 = \sigma_2 \tand \\
%                                        \nonExec{c_\data} \tand \\
%                                        w_1 = \linCons{\Phi(r_1)} \tand \\
%                                        w_2 = \linCons{\Phi(r_2)}
%                                      \end{array}
%                                      \\&\\
%                                      \sourcecolor\left.
%                                        \arraycolsep=0pt
%                                        \begin{array}[t]{rl}
%                                          \Phi' & \updReg{r_2}{0} \\
%                                                & \updReg{\pcreg}{\opc}\\
%                                                & \updReg{\rdata}{0}\\
%                                                & \updReg{\rstk}{c_\stk}\\
%                                                & \updReg{\rtmp{1}}{0}\\
%                                                & \updReg{\rtmp{2}}{0}
%                                        \end{array}\right.&
%                                      \sourcecolor\left.
%                                        \begin{array}[t]{l}
%                                          \Phi(r_1) = \retptrc(\sigma_1,\_) \tand \\
%                                          \Phi(r_2) = \retptrd(\sigma_2,(\aaddr_\stk,\eaddr_{\stk,\priv})) \tand \\
%                                          \sigma_1 = \sigma_2 \tand\\
%                                          \Phi(\rstk) = \stkptr{\rw,\stkb,\eaddr_\stk,\_} \tand \\
%                                          \Phi = (\mem,\reg,\stkf::\stk,\ms_\stk) \tand \\
%                                          \stkf = (\opc,\ms_{\stk,\priv}) \tand \\
%                                          \dom(\ms_{\stk,\priv}) = [\eaddr_\stk+1,\eaddr_{\stk,\priv}] \tand\\
%                                          \eaddr_{\stk} + 1 = \aaddr_\stk \tand\\
%                                          c_\stk = \stkptr{\rw,\stkb,\eaddr_{\stk,\priv},\aaddr_\stk} \tand\\
%                                          \Phi' = (\mem,\reg,\stk,\ms_{\stk,\priv} \uplus \ms_\stk) 
%                                        \end{array}\right.
%                                      \\
%                                      \failed & \totherwise
%                                    \end{cases}
% \end{align*}

@@

                                                        &\left[
\arraycolsep=0pt\array{l}
\reg.\rretc \mapsto \\
\hfill \retptrc(\sigma_?,(\perm,\lin,\baddr,\eaddr,\aaddr+\calllen))
\endarray
\right]\\

@@

\section{Logical Relation}
In the following definitions, \src{blue} is used to indicate values related to the source machine. This is unlike previous definitions, where \src{blue} was used to indicate source language specific parts of definitions.
\subsection{Worlds}
\[
  \World = \Worldh \times \Worlds \times \Worldfs
\]
with $\Worlds$, $\Worldh$, and $\Region$ defined as follows
\[
  \Worldh = \RegionName \parfun (\Regions + \Regionh)
\]
and
\[
  \Worlds = \RegionName \parfun (\Regions \times \SealableCaps)
\]
and
\[
  \Worldfs = \RegionName \parfun \Regions
\]
where $\RegionName = \nats$.
\[
  \Regionh = 
    \{\pure \} \times \State \times \Rels \times (\State \fun \World \monprivnefun \URel{\MemSeg^2})
\]
and
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times \State \times \Rels \times (\State \fun \World \monnefun \URel{\MemSeg^2}) \uplus \\
    \{\spatialo \} \times \State \times \Rels \times (\State \fun \World \monnefun \URel{\MemSeg^2})\uplus \\ 
    \{\revoked\}
  \end{array} \right.
\]
where $\spatial$ and $\spatialo$ are regions governing segments of memory addressed by linear capabilities. $\spatialo$ signifies that this region is addressable. $\spatial$ signifies that the region is not owned and can thus not be addressed. At the same time it signifies that if something else addresses it, it is a $\linear$ capability. Finally, $\pure$ signifies that the region is only addressed by non-linear capabilities. Notice that no region allows for both linear and non-linear capabilities to address it.

\[
  \Rels = \{(\phi_\pub, \phi) \in \powerset{\State^2}\times \powerset{\State^2} \mid \phi_\pub, \phi \text{ is reflexive and transitive and } \phi_\pub \subseteq \phi \}
\]

We introduce a bit of notation for projecting out each part of the world:
\[
  \begin{array}{l}
    \pwheap = \pi_1(W)\\
    \pwpriv = \pi_2(W)\\
    \pwfree = \pi_3(W)
  \end{array}
\]
as well as projections for the regions:
\[
  \begin{array}{l}
  \prv{(v,s,\phi_\pub,\phi,H)} = v \\
  \prv{\revoked} = \revoked     
  \end{array}
\]

We define erasure for worlds as follows:
\[
\erase{(W_{\mathrm{heap}},W_{\mathrm{priv}},W_{\mathrm{free}})}{S} =
  \left( \erase{W_{\mathrm{heap}}}{S},\erase{W_{\mathrm{priv}}}{S},\erase{W_{\mathrm{free}}}{S} \right)
\]
where erasure for each part of a world is defined as follows:
\[
\erase{W_{\mathrm{heap}}}{S} = \lambda r \ldotp \left\{
    \begin{array}{l}
      \prv{W_{\mathrm{heap}}(r)} \in S\\
      \bot
    \end{array}
  \right.
\]
\[
\erase{W_{\mathrm{priv}}}{S} = \lambda r \ldotp \left\{
    \begin{array}{l}
      \prv{\pregion{W_{\mathrm{priv}}(r)}} \in S \\
      \bot
    \end{array}
  \right.
\]
\[
\erase{W_{\mathrm{free}}}{S} = \lambda r \ldotp \left\{
    \begin{array}{l}
      \prv{W_{\mathrm{free}}(r)} \in S\\
      \bot
    \end{array}
  \right.
\]

The $\activeReg{}$ function takes a world and filters away all the revoked regions, so
\[
  \activeReg{W} = \erase{W}{\spatial,\spatialo,\pure}
\]

Disjoint union of worlds. Joins together two alike worlds with strictly disjoint ownership over $\spatialo$-regions. Two worlds can be joined together if all of their three parts agree on the region names and each of their regions can be joined together.
\[
  W_1 \oplus W_2 = W
  \text{ iff }
  \begin{array}[t]{l}
    \text{$\dom(\pwheap) = \dom(\pwheap[W_1]) = \dom(\pwheap[W_2]) \tand$} \\
    \text{$\dom(\pwfree) = \dom(\pwfree[W_1]) = \dom(\pwfree[W_2]) \tand$} \\
    \text{$\dom(\pwpriv) = \dom(\pwpriv[W_1]) = \dom(\pwpriv[W_2]) \tand$} \\
    \forall r \in \dom(\pwheap) \ldotp \pwheap(r) = \pwheap[W_1](r) \oplus \pwheap[W_2](r) \tand \\
    \forall r \in \dom(\pwfree) \ldotp \pwfree(r) = \pwfree[W_1](r) \oplus \pwfree[W_2](r) \tand \\
    \forall r \in \dom(\pwpriv) \ldotp \pi_1(\pwpriv(r)) = \pi_1(\pwpriv[W_1](r)) \oplus \pi_1(\pwpriv[W_2](r))
  \end{array}
\]
$\oplus$ on regions is defined as follows
\begin{align*}
  (\pure,\var{sm}) \oplus (\pure,\var{sm}) =  & \; (\pure,\var{sm}) \\
  (\spatial,\var{sm}) \oplus (\spatial,\var{sm}) =  & \; (\spatial,\var{sm}) \\
  (\spatialo,\var{sm}) \oplus (\spatial,\var{sm}) = & \; (\spatial,\var{sm}) \oplus (\spatialo,\var{sm})\\
                                           =  & \; (\spatialo,\var{sm})
\end{align*}
and for all other cases $\oplus$ is undefined. Specifically, $\oplus$ is not defined when both sides are a $\spatialo$-region. It is further not defined if the two sides do not agree on region type or state machine. Here $\var{sm}$ is the state machine of the region.

\subsection{Future world}
\subsubsection{\color{DarkGreen} first thoughts on change to future world relation}
{\color{DarkGreen}
The public part of the future world relation stays the same:
\begin{mathpar}
  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \pubft W.i(r) }
            { W' \future W }
\end{mathpar}
The private future world relation becomes:
\begin{mathpar}
  \inferrule{ \text{for $i \in \{\mathrm{heap},\mathrm{free},\mathrm{priv} \}$} \\ \exists m_i : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W'.i) \supseteq \dom(m_i(W.i)) \wedge \forall r \in \dom(W.i)\ldotp W'.i(m_i(r)) \privft W.i(r) \\  \exists r \in \dom(W'.\mathrm{priv}) \setminus \dom(m_{\mathrm{priv}}(\pwpriv))\ldotp \pwpriv[W'](r) \text{ is non-empty}}
            { W' \future W }
\end{mathpar}
where we take "is non-empty and stack region" to mean the following:

@@

@@

% For both $\privft$ and $\pubft$ we have
% \begin{mathpar}
%   \inferrule{ \exists m : \RegionName \fun \RegionName, \text{ injective}\ldotp \dom(W') \supseteq \dom(m(W)) \wedge \forall r \in \dom(W)\ldotp W'(m(r)) \future W(r) }
%             { W' \future W }
% \end{mathpar}
%
The private future regions allow transitions in the private part of the transition system. They also allow $\spatial$ regions to become $\revoked$:
\begin{mathpar}
  \inferrule{ }
  { \revoked \privft (\spatial,\_,\_,\_,\_)}
\and
  \inferrule{ }
            { \revoked \privft \revoked }
\and
  \inferrule{ (s,s') \in \phi }
            { (v,s',\phi_\pub,\phi,H) \privft (v,s,\phi_\pub,\phi,H)  }
\end{mathpar}
%
Public future regions are allowed to transition according to the public part of their transition system.
\begin{mathpar}
    \inferrule{ }
            { \revoked \pubft \revoked }
\and
  \inferrule{ (s,s') \in \phi_\pub }
            { (v,s',\phi_\pub,\phi,H) \pubft (v,s,\phi_\pub,\phi,H)  }
\end{mathpar}

\subsection{Memory satisfaction}
\lau{14-02-2018: No matter what the memory satisfaction ends up looking like, remember to handle the revoked case.}
{\color{DarkGreen}
Memory satisfaction for new worlds:
\[
  \memSat{\ms_S,\ms_\stk,\stk,\ms_T}{W} \text{ iff } 
  \left\{
    \begin{array}{l}
      \stk = (\opc_0,\ms_0):: \dots :: (\opc_m,\ms_m) \wedge \\
      \ms_S \uplus \ms_\stk \uplus \ms_0 \uplus \dots \uplus \ms_m  \wedge\\
      W = W_{\var{stack}} \uplus W_{\var{free\_stack}} \uplus W_{\var{heap}} \wedge\\
      \exists \ms_\var{T,stack}, \ms_\var{T,free\_stack}, \ms_\var{T,heap}\ldotp \\
      \quad \ms_T = \ms_\var{T,stack} \uplus \ms_\var{T,free\_stack} \uplus \ms_\var{T,heap} \wedge \\
      \quad \memSatStack{\stk,\ms_\var{T,stack}}{W_{\var{stack}}} \wedge \\
      \quad \memSatFStack{\ms_\stk,\ms_\var{T,free\_stack}}{W_{\var{free\_stack}}} \wedge \\
      \quad \memSatHeap{\ms_S,\ms_\var{T,heap}}{W_{\var{heap}}}
    \end{array}
  \right.
\]
